---
title: "Naive Bayes Breast Cancer"
format: html
editor: visual
---

## Dataset & Paper:

The ***Breast Cancer Wisconsin (Diagnostic)*** dataset is a renowned collection of data used extensively in machine learning and medical research. Originating from digitized images of fine needle aspirates (FNA) of breast masses, this dataset facilitates the analysis of cell nuclei characteristics to aid in the diagnosis of breast cancer. In this article, we delve into the attributes, statistics, and significance of this dataset.

[*A new classifier for breast cancer detection based on Naïve Bayesian*](https://www.sciencedirect.com/science/article/abs/pii/S0263224115002419)

### Libraries

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(e1071)
library(here)
library(ggcorrplot)
library(ggpubr)
```

### Read in Data

```{r}
breast_cancer_data <- read_csv(here("data/wisconsin_breast_cancer_data.csv"))
```

### Correlation Matrix

Let's test the assumption thatr our features are independent form each other

-   Omitted 3 columns irrelevant to correlation matrix

```{r}
breast_cancer_data_corr <- breast_cancer_data %>% 
  select(-id, -diagnosis, -...33)
```

#### Create a Matrix from Dataframe

```{r}
breast_cancer_matrix <- as.matrix(breast_cancer_data_corr)
```

#### Calculate the correlations using `cor()`

```{r}
breast_cancer_data_correlations <- cor(breast_cancer_matrix)
```

#### Plot the Correlations using `ggcorrplot`

```{r}
ggcorrplot(breast_cancer_data_correlations, method = "circle", type = "lower", lab = TRUE,  lab_size = 1)

```

#### Visualizing Correlations Between 2 Features

```{r}
ggplot(breast_cancer_data, aes(x = radius_mean, y =
                                 perimeter_mean)) + 
  geom_point() +
  geom_smooth() +
  stat_cor()
```

### Create Naive Bayes Model

```{r}
breast_cancer_mod <- naiveBayes(diagnosis ~ ., data = breast_cancer_data)
```

#### Predict Diagnosis in our Breast Cancer Dataset

```{r}
breast_cancer_data <- breast_cancer_data %>% 
  mutate(predicted_diagnosis = predict(breast_cancer_mod, 
                                       newdata = .))
```

#### Create Confusion Matrix

```{r}
breast_cancer_data %>% 
  tabyl(diagnosis, predicted_diagnosis) %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns
```
